---
title: "Network Authors"
subtitle: "Members contributing to this publication list"
---

```{r}
#| label: setup
#| include: false

library(jsonlite)
library(dplyr)
library(tidyr)
library(glue)
library(DT)
```

```{r}
#| label: load-data
#| include: false

# Read the JSON file - use relative path from project root
pub_data <- fromJSON(here::here("_data", "publications.json"))

authors <- pub_data$authors |>
  as_tibble() |>
  mutate(
    full_name = if_else(
      !is.na(display_name) & display_name != "",
      display_name,
      paste(given_name, family_name) |> trimws()
    ),
    orcid_link = glue('<a href="https://orcid.org/{orcid_id}" target="_blank">{orcid_id}</a>'),
    school_display = replace_na(school, "—")
  ) |>
  arrange(family_name, given_name)
```

## Network Members

This publication list draws from **`r nrow(authors)`** ORCID profiles. Click on any ORCID iD to view their full profile.

```{r}
#| label: author-table
#| echo: false

authors |>
  select(
    Name = full_name,
    School = school_display,
    `ORCID iD` = orcid_link
  ) |>
  datatable(
    escape = FALSE,
    filter = "top",
    options = list(
      pageLength = 50,
      order = list(list(0, "asc")),
      dom = "frtip"
    ),
    rownames = FALSE
  )
```

## Adding New Authors

To add a new author to this network publication list:

1. Ensure they have an [ORCID](https://orcid.org) profile with their publications listed
2. Add their ORCID iD to the `orcid_ids` vector in `R/fetch_orcid_publications.R`
3. The publication list will update automatically on the next scheduled run

---

*Data sourced from ORCID. Last updated: `r pub_data$last_updated`*
