---
title: "Network Publications"
subtitle: "Research outputs from our academic network"
last-updated: "Auto-updated weekly"
---

```{r}
#| label: setup
#| include: false

library(jsonlite)
library(dplyr)
library(tidyr)
library(glue)

# Load publication data
pub_data <- fromJSON(here::here("_data", "publications.json"))
```

## Overview

This website provides an automatically updated list of publications from members of our research network. Publication data is sourced directly from [ORCID](https://orcid.org) profiles and updated weekly via GitHub Actions.

Publications are automatically classified to identify those related to the **Scholarship of Teaching and Learning (SoTL)** and education research.
 
::: {.callout-note}
## Current Statistics

- **`r pub_data$n_authors`** network members
- **`r pub_data$n_publications`** unique publications
- **`r pub_data$n_sotl`** SoTL/education publications (`r round(100 * pub_data$n_sotl / pub_data$n_publications, 1)`%)
- **Last updated:** `r pub_data$last_updated`
:::

## Recent Publications

The most recent journal articles from network members:

```{r}
#| label: recent-pubs
#| echo: false
#| results: asis

pubs <- pub_data$publications |>
 as_tibble() |>
 filter(type == "journal-article") |>
 head(10)

for (i in seq_len(nrow(pubs))) {
 pub <- pubs[i, ]
 
 # Build citation string
 title_text <- if (!is.na(pub$url) && pub$url != "") {
   glue("[{pub$title}]({pub$url})")
 } else {
   pub$title
 }
 
 year_text <- if (!is.na(pub$year)) glue(" ({pub$year})") else ""
 journal_text <- if (!is.na(pub$journal) && pub$journal != "") glue(" *{pub$journal}*.") else ""
 
 cat(glue("- {title_text}{year_text}.{journal_text}\n\n"))
}
```

[View all publications →](publications.qmd)

## Quick Links

- [Browse all publications](publications.qmd) - Full searchable list
- [SoTL publications](sotl.qmd) - Education and teaching-focused research
- [Network authors](authors.qmd) - View all contributing authors
- [About this site](about.qmd) - How this works
