---
title: "SoTL Publications"
subtitle: "Scholarship of Teaching and Learning & Education Research"
---

```{r}
#| label: setup
#| include: false

library(jsonlite)
library(dplyr)
library(tidyr)
library(DT)
library(glue)
library(stringr)
library(ggplot2)
```

```{r}
#| label: load-data
#| include: false

pub_data <- fromJSON(here::here("_data", "publications.json"))

pubs <- pub_data$publications |>
  as_tibble() |>
  mutate(
    title_link = case_when(
      !is.na(url) & url != "" ~ glue('<a href="{url}" target="_blank">{title}</a>'),
      TRUE ~ title
    ),
    type_clean = str_to_title(str_replace_all(type, "-", " ")),
    year = replace_na(year, 0)
  )

sotl_pubs <- pubs |>
  filter(is_sotl == TRUE)
```

## Overview

This page highlights publications from our network that relate to the **Scholarship of Teaching and Learning (SoTL)**, educational research, and pedagogical practice. These are identified automatically based on:
 
1. **Journal matching**: Publications in known education/SoTL journals
2. **Keyword matching**: Titles containing education-related terms

::: {.callout-note}
## SoTL Statistics

- **`r nrow(sotl_pubs)`** SoTL/education publications identified
- **`r round(100 * nrow(sotl_pubs) / nrow(pubs), 1)`%** of total network output
- **`r sum(sotl_pubs$sotl_matched_on == "journal", na.rm = TRUE)`** matched by journal
- **`r sum(sotl_pubs$sotl_matched_on == "keyword", na.rm = TRUE)`** matched by keyword
:::

## SoTL Publications

```{r}
#| label: sotl-table
#| echo: false

sotl_pubs |>
  select(
    Title = title_link,
    Year = year,
    Journal = journal,
    Type = type_clean,
    `Network Authors` = network_authors,
    `Matched By` = sotl_matched_on
  ) |>
  datatable(
    escape = FALSE,
    filter = "top",
    options = list(
      pageLength = 25,
      order = list(list(1, "desc")),
      columnDefs = list(
        list(width = "35%", targets = 0),
        list(width = "8%", targets = 1),
        list(className = "dt-left", targets = "_all")
      ),
      dom = "Bfrtip",
      buttons = c("csv", "excel")
    ),
    extensions = "Buttons",
    rownames = FALSE
  )
```

## SoTL Publications by Year

```{r}
#| label: sotl-by-year
#| echo: false
#| fig-width: 10
#| fig-height: 5

year_counts <- sotl_pubs |>
  filter(year > 0) |>
  count(year) |>
  arrange(year)

if (nrow(year_counts) > 0) {
  ggplot(year_counts, aes(x = year, y = n)) +
    geom_col(fill = "#18bc9c") +
    labs(
      x = "Year",
      y = "Number of Publications",
      title = "SoTL Publications per Year"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
    )
} else {
  cat("No publications with valid years to display.")
}
```

## Top Journals

```{r}
#| label: top-journals
#| echo: false
#| results: asis

journal_counts <- sotl_pubs |>
  filter(!is.na(journal), journal != "") |>
  count(journal, sort = TRUE) |>
  head(15)

if (nrow(journal_counts) > 0) {
  for (i in seq_len(nrow(journal_counts))) {
    cat(glue("- **{journal_counts$journal[i]}**: {journal_counts$n[i]}\n"))
  }
} else {
  cat("No journal information available.")
}
```

## Classification Method

Publications are classified as SoTL/education-related using a two-stage process:

### Stage 1: Journal Matching

If a publication appears in a known education or SoTL journal, it is automatically classified. The current list includes `r length(pub_data$sotl_journals)` journals across:

- General higher education journals
- Medical and health professions education
- Science education (chemistry, biology, physics)
- Discipline-specific pedagogy journals
- Educational technology journals

### Stage 2: Keyword Matching

For publications not matched by journal, titles are scanned for `r length(pub_data$sotl_keywords)` education-related keywords and phrases, including:

- Core terms: teaching, learning, pedagogy, curriculum, assessment
- Higher education: undergraduate, postgraduate, transitions, widening participation
- Methods: flipped classroom, problem-based learning, simulation
- Professional education: clinical teaching, placements, interprofessional

::: {.callout-warning}
## Limitations

This automated classification is imperfect. Some education-related publications may be missed (false negatives), and some non-education publications may be incorrectly included (false positives). The keyword and journal lists can be expanded by editing `R/fetch_orcid_publications.R`.
:::

---

*Data sourced from ORCID. Last updated: `r pub_data$last_updated`*
