---
title: "Themes"
---

```{r}
#| label: setup
#| include: false

library(jsonlite)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(glue)
library(DT)
library(ggplot2)

themes_path <- here::here("_data", "themes.json")

# jsonlite silently treats a non-existent path as literal JSON text, which fails
# with an unhelpful lexical error. Check first and say what to do about it.
if (!file.exists(themes_path)) {
  stop(
    "\n_data/themes.json not found.\n",
    "Generate it before rendering:\n",
    "  Rscript R/fetch_abstracts.R    # abstracts, cached; a few minutes first run\n",
    "  Rscript R/classify_themes.R   # writes _data/themes.json\n",
    call. = FALSE
  )
}

theme_data <- fromJSON(themes_path, simplifyVector = TRUE)

# A themes.json written before the publication filter existed carries every
# output type, so the chart and table would quietly show inflated counts.
if (is.null(theme_data$publication_types)) {
  stop(
    "\n_data/themes.json predates the publication-type filter.\n",
    "Rebuild it:\n",
    "  Rscript R/classify_themes.R\n",
    call. = FALSE
  )
}

theme_defs <- as_tibble(theme_data$themes)
counts     <- as_tibble(theme_data$counts)

# Describe what is being counted, driven by the classifier's own configuration
# so this prose cannot drift away from publication_types.
type_label <- function(x) {
  lookup <- c(
    "journal-article" = "journal articles",
    "book"            = "books",
    "book-chapter"    = "book chapters",
    "preprint"        = "preprints",
    "report"          = "reports"
  )
  out <- unname(lookup[x])
  fallback <- paste0(str_replace_all(x, "-", " "), "s")
  if_else(is.na(out), fallback, out)
}

#' Join with commas and a final "and" (avoids a stringr >= 1.5 dependency)
comma_and <- function(x) {
  n <- length(x)
  if (n == 0) return("")
  if (n == 1) return(x)
  paste0(paste(x[-n], collapse = ", "), " and ", x[n])
}

included_types <- comma_and(type_label(theme_data$publication_types))

outputs <- as_tibble(theme_data$outputs) |>
  mutate(
    theme_list = str_split(replace_na(themes, ""), ";"),
    theme_list = map(theme_list, ~ .x[.x != ""]),
    n_themes   = lengths(theme_list)
  )

# UofG brand colours
uofg_blue <- "#003865"
```

This page counts **`r included_types` only**. `r theme_data$n_outputs` publications by network members are flagged as Scholarship of Teaching and Learning and sorted into themes below, so you can see what work is happening where.

A further `r theme_data$n_excluded` SoTL outputs are excluded — conference papers, abstracts, posters and presentations, preprints, reports, blog posts and theses. They appear on the [All Outputs](publications.qmd) page. Themes were revised after the network symposium in August 2026.

::: {.callout-warning collapse="true"}
## How this grouping is made, and where it goes wrong

The grouping is automatic. Each output is matched on its title, journal name and — where one could be retrieved — its abstract, against a set of keywords for each theme. Nothing here has been read by a person.

- **A publication can appear under more than one theme.** Most SoTL work touches several at once, and forcing a single label would misrepresent it. `r sum(outputs$n_themes > 1)` carry more than one theme.
- **`r theme_data$n_unclassified` are not matched to any theme.** Usually the title is too oblique for the matching to work, not that the work sits outside every theme.
- **Some are in the wrong place.** Abstracts were available for `r theme_data$n_with_abstract` of `r theme_data$n_outputs`; the rest were matched on title and journal alone, which is thin evidence.
:::

## Publications per theme

```{r}
#| label: theme-chart
#| echo: false
#| fig-width: 9
#| fig-height: 4.5

plot_counts <- counts |>
  mutate(label = str_wrap(label, 28)) |>
  arrange(n)

ggplot(plot_counts, aes(x = reorder(label, n), y = n)) +
  geom_col(fill = uofg_blue, width = 0.7) +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5, colour = uofg_blue) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    x = NULL,
    y = glue("Number of publications ({included_types})")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.title         = element_text(colour = uofg_blue),
    axis.text.y        = element_text(colour = uofg_blue, lineheight = 0.95)
  )
```

Each bar counts the `r theme_data$n_outputs` publications on this page, not all `r theme_data$n_outputs + theme_data$n_excluded` SoTL outputs. Bars sum to more than `r theme_data$n_outputs` because a publication can sit in several themes. The differences between themes reflect what has been published and registered on ORCID.

## Browse by theme

```{r}
#| label: master-table
#| echo: false

label_lookup <- set_names(theme_defs$label, theme_defs$id)

# One row per publication-theme pair. DT only renders a dropdown filter for
# factor columns, and it matches the whole cell exactly, so a combined "A; B"
# cell could never be found by picking "A". Exploding keeps the filter usable;
# a publication in several themes appears once under each.
table_data <- outputs |>
  mutate(
    Title = if_else(
      !is.na(url) & url != "",
      glue('<a href="{url}" target="_blank">{title}</a>'),
      title
    ),
    Theme = map(theme_list, function(ids) {
      if (length(ids) == 0) return("Unclassified")
      unname(label_lookup[ids])
    })
  ) |>
  unnest_longer(Theme) |>
  # Factor levels fix the dropdown order and keep Unclassified last
  mutate(Theme = factor(Theme, levels = c(theme_defs$label, "Unclassified"))) |>
  select(Theme, Title, Year = year, Journal = journal,
         `Network authors` = network_authors) |>
  arrange(Theme, desc(Year))

datatable(
  table_data,
  escape  = FALSE,
  filter  = "top",
  options = list(
    pageLength = 25,
    order      = list(list(2, "desc")),
    columnDefs = list(
      list(width = "22%", targets = 0),
      list(width = "32%", targets = 1),
      list(width = "5%",  targets = 2),
      list(className = "dt-left", targets = "_all")
    )
  ),
  rownames = FALSE
)
```

Pick a theme from the dropdown at the top of the first column. The table covers the same `r theme_data$n_outputs` publications as the chart; one spanning several themes is listed once under each, so there are more rows than publications.

------------------------------------------------------------------------

*Themes assigned automatically from title, journal and abstract. Last updated: `r theme_data$last_updated`*
